
  
```{r}
library('lattice') 
library("plyr") 
library('np')
```

#### load data
```{r}
training_data <- read.delim('geno_train.txt', sep=',', header=T)
testing_data <- read.delim('geno_test.txt', sep=',', header=T)
```


## (a)
### Plotting the 15 plots, I could manage it with par(mfrow=c(4,5), mar = c(2,2,1,1)) 
```{r}
par(mfrow=c(4,5), mar = c(2,2,1,1))
plot(density(training_data[training_data$population=='Hispanic',]$D8S1179), main='D8S1179', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D8S1179), col='red')
lines(density(training_data[training_data$population=='African American',]$D8S1179), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D21S11), main='D21S11', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D21S11), col='red')
lines(density(training_data[training_data$population=='African American',]$D21S11), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D7S820), main='D7S820', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D7S820), col='red')
lines(density(training_data[training_data$population=='African American',]$D7S820), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$CSF1PO), main='CSF1PO', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$CSF1PO), col='red')
lines(density(training_data[training_data$population=='African American',]$CSF1PO), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D3S1358), main='D3S1358', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D3S1358), col='red')
lines(density(training_data[training_data$population=='African American',]$D3S1358), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$TH01), main='TH01', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$TH01), col='red')
lines(density(training_data[training_data$population=='African American',]$TH01), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D13S317), main='D13S317', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D13S317), col='red')
lines(density(training_data[training_data$population=='African American',]$D13S317), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D16S539), main='D16S539', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D16S539), col='red')
lines(density(training_data[training_data$population=='African American',]$D16S539), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D2S1338), main='D2S1338', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D2S1338), col='red')
lines(density(training_data[training_data$population=='African American',]$D2S1338), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D19S433), main='D19S433', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D19S433), col='red')
lines(density(training_data[training_data$population=='African American',]$D19S433), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$vWA), main='vWA', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$vWA), col='red')
lines(density(training_data[training_data$population=='African American',]$vWA), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$TPOX), main='TPOX', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$TPOX), col='red')
lines(density(training_data[training_data$population=='African American',]$TPOX), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D18S51), main='D18S51', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D18S51), col='red')
lines(density(training_data[training_data$population=='African American',]$D18S51), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$D5S818), main='D5S818', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$D5S818), col='red')
lines(density(training_data[training_data$population=='African American',]$D5S818), col='blue')

plot(density(training_data[training_data$population=='Hispanic',]$FGA), main='FGA', col='green')
lines(density(training_data[training_data$population=='Caucasian',]$FGA), col='red')
lines(density(training_data[training_data$population=='African American',]$FGA), col='blue')


plot(5, type = "n", axes=FALSE, xlab="", ylab="")
legend(x = "top",inset = 0,
        legend = c("Hispanic", "Caucasian", "African American"), col=c('green', 'red', 'blue'), lwd=5, cex=.5, horiz = FALSE)

```



## (b)

##### LDA
```{r}
library('MASS')
model_lda <- lda(population ~ ., data=training_data)
pred_lda <- predict(model_lda, testing_data[,-16])$class
```

#### Printing the model summary
```{r}
model_lda
```

### LDA Misclassification rate
```{r}
lda_misclassification <- (sum(pred_lda!=testing_data$population)/ length(testing_data$population))*100 
lda_misclassification
```
### We get misclassification of 34.911 %

##### SVM for classification
```{r}
library('e1071')
training_data$population<-as.factor(training_data$population)
svm_classfication <- svm(population~ ., data=training_data, scale = TRUE,type = "C-classification")
pred_svm_classfication <- predict(svm_classfication, testing_data[,-16])
```

### Model Summary
```{r}
summary(svm_classfication)
```
### Misclassification
```{r}
svm_misclassification <- (sum(pred_svm_classfication!=testing_data$population)/ length(testing_data$population))*100 
svm_misclassification
```
### We get misclassification of 30.1775 %

## (c)

##### PCA
```{r}
gene_pr_model <- prcomp(training_data[,-16], center = TRUE, scale = TRUE)
summary(gene_pr_model)
```


#### In our dataset we have our top three Principal components can explain 26.8% of the variance in the data. We will go with that.

```{r}
gene_pc_mod <- data.frame(gene_pr_model$x[,1:3])
gene_pc_mod <- cbind(gene_pc_mod, training_data$population)
colnames(gene_pc_mod)[4] <- "population"
```


```{r}
pca_testing <- data.frame(predict(gene_pr_model, testing_data[,-16])[, 1:3])
```

```{r}
pc_lda_model <- lda(population ~ ., data = gene_pc_mod)
pc_lda_predictions <- predict(pc_lda_model, newdata = pca_testing)$class
```


```{r}
pc_lda_mis <- (sum(pc_lda_predictions!=testing_data$population)/ length(testing_data$population))*100 
pc_lda_mis
```
### We get misclassification of 33.72781 %

```{r}
svm_model_pc <- svm(population~ ., data=gene_pc_mod)
svm_pred <- predict(svm_model_pc, pca_testing)
```


```{r}
pc_svm_mis <- (sum(svm_pred!=testing_data$population)/ length(testing_data$population))*100 
pc_svm_mis
```
### We get misclassification of 32.54438 %






